---
title: "IS605_HW12"
author: "Daina Bouquin"
date: "April 24, 2016"
output: html_document
---

```{r}
library(stats)
library(boot)
```

Using the stats and boot libraries in R perform a cross-validation experiment to observe
the bias variance tradeoff. You’ll use the auto data set from previous assignments. This
dataset has 392 observations across 5 variables. We want to fit a polynomial model of
various degrees using the glm function in R and then measure the cross validation error
using cv.glm function.
```{r}
# download and read in the auto data
download.file("https://raw.githubusercontent.com/dbouquin/IS_605/master/IS605_HW5/auto-mpg.data", "auto-mpg.data", method="curl")
auto_data <- read.table('auto-mpg.data', 
                        col.names=c('disp', 'hp', 'wt', 'acc', 'mpg'))
str(auto_data)
```
Fit various polynomial models to compute mpg as a function of the other four variables
acc, wt, hp, and disp using glm function.
```{r}
glm_fit1 <- glm(mpg~poly(disp+hp+wt+acc,1), data=auto_data)
glm_fit2 <- glm(mpg~poly(disp+hp+wt+acc,2), data=auto_data)
glm_fit3 <- glm(mpg~poly(disp+hp+wt+acc,3), data=auto_data)
glm_fit4 <- glm(mpg~poly(disp+hp+wt+acc,4), data=auto_data)
glm_fit5 <- glm(mpg~poly(disp+hp+wt+acc,5), data=auto_data)
glm_fit6 <- glm(mpg~poly(disp+hp+wt+acc,6), data=auto_data)
glm_fit7 <- glm(mpg~poly(disp+hp+wt+acc,7), data=auto_data)
glm_fit8 <- glm(mpg~poly(disp+hp+wt+acc,8), data=auto_data)

set.seed(250)
# use cv.glm() on the results
# ?cv.glm
cv_err1 <- cv.glm(auto_data,glm_fit1,K=5)$delta[1]
cv_err2 <- cv.glm(auto_data,glm_fit2,K=5)$delta[1]
cv_err3 <- cv.glm(auto_data,glm_fit3,K=5)$delta[1]
cv_err4 <- cv.glm(auto_data,glm_fit4,K=5)$delta[1]
cv_err5 <- cv.glm(auto_data,glm_fit5,K=5)$delta[1]
cv_err6 <- cv.glm(auto_data,glm_fit6,K=5)$delta[1]
cv_err7 <- cv.glm(auto_data,glm_fit7,K=5)$delta[1]
cv_err8 <- cv.glm(auto_data,glm_fit8,K=5)$delta[1]

# combine the cv.glm results
cv_err <- rbind(cv_err1, cv_err2,cv_err3,cv_err4,cv_err5,cv_err6,cv_err7,cv_err8)

# plot the resulting 5 fold cross-validation curve to see the the tradeoff between bias and variance.
plot(x = 1:8, cv_err, type='b')
```
Feedback:

Should have been done with a loop 
e.g.

for (deg in 1:8) {

glm.fit=glm(mpg~poly(displacement+horsepower+weight+acceleration, deg), data=auto_mpg)

... and then do the cross-validation
