---
title: "IS605_HW2"
author: "Daina Bouquin"
date: "February 14, 2016"
output: html_document
---
This file can be seen fully rendered on RPubs [here]().

####Problem Set 1:
#####1. Show that $A^TA\neq AA^T$  
```{r}
# First create matrix A
A <- matrix(c(8,9,10,0,1,2,11,12,13), nrow = 3)
A

# Then transpose matrix A using t()
t_A <- t(A)
t_A

# Multiply the transposed matrix A by matrix A
ATA <- t_A %*% A
ATA

# Multiply A by transposed matrix A (AA^T)
AAT <- A %*% t_A
AAT
```
This shows that generally $A^TA\neq AA^T$. You can check this using logical equality test:
```{r}
# Check using logical comparison
ATA == AAT
```
#####2. For a special type of square matrix $A$, we get $A^TA = AA^T$. Under what conditions could this be true?

For the identity matrix, its multiples, and symetric matrices $A^TA = AA^T$. This is true because a symmetric matrix and its transpose are equal, meaning you are just multiplying a matrix by itself.     

```{r}
# Show this using 2x2 identiy matrix
id <- matrix(c(1,0,0,1), nrow = 2)
id
# Transpose the id matrix
t_id <- t(id)
# try both
id1 <- t_id %*% id
id2 <- id %*% t_id
# check using logic
id1 == id2
```

#####3. Write an R function to factorize a square matrix $A$ into LU or LDU 
This function will factorize a 2x2 and a 3x3 matrix into LU
```{r}
f <- function(A) {

#Check dimensions of the matrix - 2x2 or 3x3

# 2x2
if (nrow(A) == 2) {

  #Subract multipler
  E <- -A[2,1] / A[1,1]
  L <- matrix(c(1,0,E,1),nrow=2)
  U <- L %*% A

  L <- solve(L)
  
  print ("------ A ------")
  print (A)
  print ("------ L ------")
  print(L)
  print ("------ U ------")
  print(U)

  (L %*% U == A)

#3x3
} else if (nrow(A) == 3) {

  #Subract multipler
  e_21 <- -A[2,1] / A[1,1]
  me_21 <- matrix(c(1,0,0,e_21,1,0,0,0,1), nrow=3)
  A2 <- me_21 %*% A

  #Subract multipler
  e_31 <- -A2[3,1] / A2[1,1]
  me_31 <- matrix(c(1,0,0,0,1,0,e_31,0,1), nrow=3)
  A3 <- me_31 %*% A2

  #Subtract multipler
  e_32 <- -A3[3,2] / A3[2,2]
  me_32 <- matrix(c(1,0,0,0,1,0,0,e_32,1), nrow=3)
  U <- me_32 %*% A3

  L <- solve(me_21) %*% solve(me_31) %*% solve(me_32)

  A <- L %*% U
  
  print ("------ A ------")
  print(A)
  print ("------ L ------")
  print(L)
  print ("------ U ------")
  print(U)

  (L %*% U == A)
 
  }
}
```

Check with a 2x2 and a 3x3 matrix
```{r}
# 2x2
A2x2 <- matrix(c(2,3,4,5), nrow = 2)
f(A2x2)

# 3x3
A3x3 <- matrix(c(8,9,10,0,1,2,11,12,13), nrow = 3)
f(A3x3)
```

