---
title: "IS605_HW1"
author: "Daina Bouquin"
date: "February 6, 2016"
output: html_document
---
    
####Problem Set 1:
#####1. Calculate the dot product of $u.v$ where $u=[0.5;0.5]$ and $v=[3;-4]$  
```{r}
u= c(0.5,0.5)
v=c(3,-4)
dp_uv<- u%*%v # %*% = matrix multiplication. If both vectors same length inner-product is returned
dp_uv
```
#####2. What are the lengths of $u$ and $v$? 
```{r}
# Euclidean norm of the vector = the square root of the inner-product of a vector with itself.
dp_uu <- u%*%u
dp_vv <- v%*%v
```
Length of u:
```{r}
l_u <- sqrt(dp_uu)
l_u
```
Length of v:
```{r}
l_v <- sqrt(dp_vv)
l_v
```
#####3. What is the linear combination $3u-2v$?
```{r}
(3*u)-(2*v)
```
#####4. What is the angle between $u$ and $v$?
```{r}
#The dot-product between two vectors is proportional to their lengths and angle between them.
#dp_uv = cosine of the angle between u and v
theta <- acos(dp_uv/(l_u*l_v)) # acos() = arc-cosine
theta
```
   
####Problem Set 2:
Set up a system of equations with 3 variables and 3 constraints and solve for x.  Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination.  Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. Please note that you do have to worry about zero pivots, and that you should not use the built-in function `solve` to solve this system or use matrix inverses.The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.      
   
Please test it with the system below and it should produce a solution $x = [-1.55;-0.32;0.95]$
$$\begin{bmatrix}1 & 1 & 3 \\2 & -1 & 5 \\-1 & -2 & 4\end{bmatrix}\,\begin{bmatrix}x_1 \\x_2 \\x_3\end{bmatrix}=\begin{bmatrix}1 \\2 \\6\end{bmatrix}$$   
   
```{r}


```

