---
title: "IS605_HW1"
author: "Daina Bouquin"
date: "February 6, 2016"
output: html_document
---
This file can be seen fully rendered on RPubs [here](http://rpubs.com/dbouquin/IS605_HW1).
   
####Problem Set 1:
#####1. Calculate the dot product of $u.v$ where $u=[0.5;0.5]$ and $v=[3;-4]$  
```{r}
u= c(0.5,0.5)
v=c(3,-4)
dp_uv<- u%*%v # %*% = matrix multiplication. If both vectors same length inner-product is returned
dp_uv
```
#####2. What are the lengths of $u$ and $v$? 
```{r}
# Euclidean norm of the vector = the square root of the inner-product of a vector with itself.
dp_uu <- u%*%u
dp_vv <- v%*%v
```
Length of u:
```{r}
l_u <- sqrt(dp_uu)
l_u
```
Length of v:
```{r}
l_v <- sqrt(dp_vv)
l_v
```
#####3. What is the linear combination $3u-2v$?
```{r}
(3*u)-(2*v)
```
#####4. What is the angle between $u$ and $v$?
```{r}
#The dot-product between two vectors is proportional to their lengths and angle between them.
#dp_uv = cosine of the angle between u and v
theta <- acos(dp_uv/(l_u*l_v)) # acos() = arc-cosine
theta # given in radians
```
   
####Problem Set 2:
Set up a system of equations with 3 variables and 3 constraints and solve for x.Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination. Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. Please note that you do have to worry about zero pivots, and that you should not use the built-in function `solve` to solve this system or use matrix inverses.The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.      
   
Please test it with the system below and it should produce a solution $x = [-1.55;-0.32;0.95]$
$$\begin{bmatrix}1 & 1 & 3 \\2 & -1 & 5 \\-1 & -2 & 4\end{bmatrix}\,\begin{bmatrix}x_1 \\x_2 \\x_3\end{bmatrix}=\begin{bmatrix}1 \\2 \\6\end{bmatrix}$$   

####### The below function will only work for 3-equation systems 
```{r}
#inputs to sys_solver function are matrix A & constraint vector b

A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim=c(3,3))
A # Make sure it looks like our problem prompt
b <- c(1, 2, 6) # constraint vector

sys_solver <- function(A, b) {
  # create a vector to store outputs.
  x = numeric(3) # size of the vector
  # bind A and b to create a new matrix (augmented matrix)
  new_m = cbind(A, b)
  
  # need 1 in top left pivot
  if (new_m[1,1] == 0) {
    #  need to switch with row that has non-zero in first column
    if (new_m[2,2] != 0) {
    # switch rows 1 and 2 as row 2 has a non-zero in column 1
    new_m = new_m[c(2,1,3),]
    } else {
    # switch rows 3 and 1 since row 3 has a non-zero
      new_m = new_m[c(3,2,1),]
    }
  }  else if (new_m[1,1] != 1) {
    # else divide 1st row by row 1 column 1 to get 1 in pivot
    new_m[1,] <- new_m[1,] / new_m[1,1]
  }  

  # Zero other entries in column 1
  if (new_m[2,1] !=0) {
    # use value of row 2 column 1 for vector addition to zero the row
    vec <- new_m[2,1] * new_m[1,]
    # reduce row 2 using the new vec object
    new_m[2,] <- -new_m[2,] + vec
  }
  if (new_m[3,1] !=0) {
    # use value of row 3 column 1 for vector addition to zero second row
    vec <- new_m[3,1] * new_m[1,]
    # now reduce row 2 using the vec
    new_m[3,] <- -new_m[3,] + vec
  }
  
  # check if row 2 column 2 are 0. If no, switch for row 3 to get non-zero
  if (new_m[2,2] == 0) {
    # switch rows 2 and 3 as row 2 has 0 in column 2
    new_m = new_m[c(1,3,2),]
  } 
  
  # divide second row by row 2 column 2 to get 1 in pivot
  new_m[2,] <- new_m[2,] / new_m[2,2]

  # use value of row 3 column 2 for vector addition to zero row 3 column 2
  vec <- new_m[3,2] * new_m[2,]
  # now reduce row 3 using tempvec
  new_m[3,] <- -new_m[3,] + vec  

  # solve for x with back subsitution
  x[3] <- new_m[3,4] / new_m[3,3]
  x[2] <- (new_m[2,4] - new_m[2,3]*x[3]) / new_m[2,2]
  x[1] <- (new_m[1,4] - new_m[1,2]*x[2] - new_m[1,3]*x[3]) / new_m[1,1]

  return(x)
}

# Try it out!

sys_solver(A,b)
```

####### Check our work using the built in `solve` function:
```{r}
# First will show the solution using built in solve() so we can check the answer:
A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim=c(3,3))
b <- c(1, 2, 6)
solve(A, b)
```