---
title: "IS605_HW7"
author: "Daina Bouquin"
date: "March 16, 2016"
output: html_document
---

1. Please write a function to compute the expected value and standard deviation of an array of values. Compare your results with that of R’s mean and std functions. 
```{r}

# create a small example dataset
set.seed(20)
x <- rnorm(n=20, mean=2, sd=1.5)
x
n <- length(x) 
  
# calculates both mean and sd of x
Mean_SD <- function(x) 
{
  man_mean <- sum(x)/n

  man_sd <- sqrt(sum((x - mean(x))^2) / (n - 1))

  return (c("Mean" = man_mean, "SD" = man_sd))

}

Mean_SD(x)
mean(x)
sd(x)
```

2. Now, consider that instead of being able to neatly fit the values in memory in an array,you have an infinite stream of numbers coming by. How would you estimate the mean and standard deviation of such a stream?
```{r}
# create a file that can be continually updated and doesn't have to be stored constantly, it'll just be read as needed with new data if it's there.

create_temp_file <- function(temp_name)
{
  # create empty numeric variables to store data later
   length <- 0
   sum <- 0
   sum_sq <- 0
   mean <- 0
   sd <- 0
   # write the empty variables to a df and then to csv
   temp_file <- data.frame(length, sum, sum_sq, mean, sd)
   write.csv(temp_file, temp_name)
}

# Calculate the mean and sd of the data found in the file
moving_mean <- function(new_data)
  # first check if the file is in the working directory, if not create one.
{
  fileName <- "tempfile.csv"
  if (file.exists(fileName))
  {
    temp_file <- read.csv(fileName)
  }
  else
  {
    create_temp_file(fileName)
  }

  # Map the current value colnames to variables
  # If this isn't done there is an error: differing number of rows
  length_a  <- temp_file$length
  sum_a   <- temp_file$sum
  sum_sq_a <- temp_file$sum_sq

  # Calculate based ond data found in temp_file
  length_b <- length_a + 1
  sum_b <- sum_a + new_data
  sum_sq_b <- sum_sq_a + new_data^2 

  # Find mean and sd
  mean_b <- sum_b/length_b
  sd_b <- sqrt((length_b)*sum_sq_b-sum_b^2)/length_b

  # create a new variable to be written over past temp_file
  temp_file_new <- data.frame(length = length_b, sum = sum_b, sum_sq = sum_sq_b, mean=mean_b, sd=sd_b)
  write.csv(temp_file_new, fileName)

  # return the calculations
  return (c("mean" = mean_b, "sd" = sd_b))

}

```
Give the functions a try with a few sample datasets
```{r}
## Test 1 ##
set.seed(10)
test <- rnorm(n=500, mean=25, sd=1.5)

create_temp_file("tempfile.csv")

for(i in test)
{
    mean_rolling <- moving_mean(i)
}

mean_rolling

## Test 2 ##
set.seed(20)
test2 <- rnorm(n=1000, mean=55, sd=1)

create_temp_file("tempfile.csv")

for(i in test2)
{
    mean_rolling2 <- moving_mean(i)
}

mean_rolling2
```


