---
title: "IS605_HW13"
author: "Daina Bouquin"
date: "May 4, 2016"
output: html_document
---

1. Write a program to compute the derivative of $f(x) = x^3 + 2x^2$ at any value of $x$. Your function should take in a value of $x$ and return back an approximation to the derivative of $f(x)$ evaluated at that value. You should not use the analytical form of the derivative to compute it. Instead, you should compute this approximation using limits.
```{r}
# define the function
f <- function(x) {x^3 + 2*x^2}

# use limits to compute derivative
der <- function(x) {
  h = 0.00000000001
  der = (f(x + h) - f(x))/(h)
  return(der) # more appropriate to round? (approximation)
}

# try out the function on some values of x
der(3)
der(10)
```

2. Now, write a program to compute the area under the curve for the function $3x^2+4x$ in the range $x = [1, 3]$. You should first split the range into many small intervals using some really small $∆x$ value (say 1e-6) and then compute the approximation to the area under the curve.
```{r}
# split the range into many small intervals using 1e-6
x <- seq(1, 3, by=1e-6)
# represent the function
f <- 3*(x^2) + 4*x
# approx areas
areas <- f*(1e-6)
sum(areas)
````

3. Please solve the remaining problems analytically    
    
* Use integration by parts to solve for $\int sin(x)cos(x)dx$. MathWorld refresher [here](http://mathworld.wolfram.com/IntegrationbyParts.html).   
$d(uv)=udv+vdu$         
$u = sin(x),  du = cos(x) dx$  
apply the power rule    
$\frac{u^{1+1}}{1+1}$   
substitute back $u= sin(x)$
$\frac{sin^{1+1}(x)}{1+1}$ ∴   
$$\int sin(x)cos(x)dx = \frac{sin^2(x)}{2}+C$$    

* Use integration by parts to solve for $\int x^2e^x dx$   
$\int uv' = uv - \int u'v$         
$u = x^2,  u'=2x,  v'= e^x,  v= e^x$   
$= x^2e^x - \int 2xe^x dx$   
$= \int 2xe^x dx = 2(e^xx-e^x)$    
$= x^2e^x-2(e^xx-e^x)$ ∴   
$$\int x^2e^x dx = x^2e^x-2(e^xx-e^x)+C$$
   

* What is $\frac{d}{dx}(x cos(x))$ ?    
$\frac{d}{dx}(x)= 1$    
$\frac{d}{dx}(cos(x)) = -sin(x)$    
$= 1cos(x) + (-sin(x))x$ ∴   
$$\frac{d}{dx}(x cos(x)) = cos(x)-xsin(x)$$
    
* What is $\frac{d}{dx}(e^{x^4})$ ?   
chain rule: $\frac{df(u)}{dx} = \frac{df}{du}\cdot\frac{du}{dx}$   
$x^4 = u$
$=\frac{d}{du}(e^u)\frac{d}{dx}(x^4)$   
$\frac{d}{du}(e^u) = e^u$   
$\frac{d}{dx}(x^4) = 4x^3$   
$= e^u4x^3$    
sub back in $u = x^4$ ∴     
$$\frac{d}{dx}(e^{x^4}) = e^{x^4}4x^3$$