---
title: "IS605_HW10"
author: "Daina Bouquin"
date: "April 7, 2016"
output: html_document
---

[PageRank assignment](https://github.com/dbouquin/IS_605/blob/master/IS605_HW10.pdf)

In the original formulation of the algorithm, the PageRank of a page $P_i$, denoted by
$r(P_i)$, is the sum of the PageRanks of all pages pointing into $P_i$. Where $B_{P_i }$ is the set of pages pointing into $P_i$ and $|P_i|$ is the number of outlinks from $P_i$.

$$ r(P_{ i })=\sum _{ P_{j∈}B_{P_i }}\frac{r(P_{ j })}{|P_{ j }|} $$

You’ll verify for yourself that PageRank works by performing calculations on a small
universe of web pages. Let’s use the 6 page universe that we had in the course notes. For this directed graph, perform the following calculations in R.
     
#####1. Form the $A$ matrix. Then, introduce decay and form the $B$ matrix as we did in [the course notes](https://github.com/dbouquin/IS_605/blob/master/wk10_PageRank.pdf).

"The $A$ matrix is column normalized such that each element. $A_{ji} = \frac{1}{|P_i|}$ if there is a link from page $i$ to page $j$ and 0 otherwise. Note that this can be interpreted as the probability of clicking one of the links in page $i$ at random. 
A new matrix $B$ which is a decayed version of the original matrix $A$ is defined as $B = 0.85 \times A + \frac{0.15}{n}$. To make every row add up to unity, the uniform probability is properly adjusted to be $(1 − d)$ where $d$ is the decay applied to the original transition matrix A. In the classic version by Page and Brin, they chose $d = 0.85$"

```{r}
# A from prompt
c1 <- c(0,0.5,0.5,0,0,0)
c2 <- c(0,0,1,0,0,0)
c3 <- c(0.25,0.25,0,0,0.25,0.25)
c4 <- c(0,0,0,0,0.5,0.5)
c5 <- c(0,0,0,0.5,0,0.5)
c6 <- c(0,0,0.5,0.5,0,0)

A <-as.matrix(cbind(c1, c2, c3, c4, c5, c6))
A

B <- 0.85*A + 0.15/6 # n = number of pages ; d = 0.85
B
```
#####2. Start with a uniform rank vector $r$ and perform power iterations on B until convergence. That is, compute the solution $r = Bn \times r$. Attempt this for a sufficiently large $n$ so that $r$ actually converges.
```{r}
# create a uniform rank vector
r <- rep(1/nrow(A), nrow(A)) # 1/number of rows
r

# perform power iterations on B
# Function takes X matrix as input
power_it <- function(X, r, n) {
  it = diag(nrow(X)) # diag() construct or replace a diagnonal matrix
  # loop through n
  for (i in 1:n)
  {
    it = it %*% X
  }
  return (it %*% r)
}

# try out the function on different n
power_it(B, r, 10)
power_it(B, r, 20)
power_it(B, r, 30) #Stable!
power_it(B, r, 40)
power_it(B, r, 50)
```
#####3. Compute the eigen-decomposition of B and verify that you indeed get an eigenvalue of 1 as the largest eigenvalue and that its corresponding eigenvector is the same vector that you obtained in the previous power iteration method. Further, this eigenvector has all positive entries and it sums to 1.
```{r, warning=FALSE}
e_B <- eigen(B)
ev_B <- as.numeric(e_B$values)
ev_B # greatest value == 1
e_B$vectors[,1]

# check to make sure the answer is the same as power iteration method
check <- as.numeric((1/sum(e_B$vectors[,1]))*e_B$vectors[,1])
check
stable <- power_it(B, r, 30) # this was where we got stability (convergence)
# without 'rounding' to the 7th digit, the result is FALSE
round(check,7) == round(stable,7)

# you can see that the values are all positive and they sum to 1
sum(check) == 1

```
#####4. Use the `igraph` package (`graph` is depreciated?) in R and its `page.rank` method to compute the Page Rank of the graph as given in A. Note that you don’t need to apply decay. The package starts with a connected graph and applies decay internally. Verify that you do get the same PageRank vector as the two approaches above.
```{r, warning=FALSE, message=FALSE}
library(igraph)

#?graph.adjacency()
graph_A <- graph.adjacency(t(A), weighted = TRUE, mode = "directed") # must transpose, weight (default == NULL)
# page.rank() == Google PageRank
pr_graph_A <- page.rank(graph_A)$vector

# make sure that the stable page rank calculated first are the same
# without 'rounding' to the 7th digit, the result is FALSE
round(pr_graph_A,7) == round(stable,7)
```